---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load the data and transform date factor to proper format:

```{r}
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?

```{r}
stepsPerDay <- aggregate(steps~date, data = activity, sum)
hist(stepsPerDay$steps)
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)
```
The average number of steps per day = `r meanSteps`  
The median number of steps per day = `r medianSteps`  


## What is the average daily activity pattern?

```{r}
stepsPerInterval <- aggregate(steps~interval, data = activity, mean)
plot(stepsPerInterval, type = "l")
maxStepsInterval <- stepsPerInterval[stepsPerInterval$steps == max(stepsPerInterval$steps), 1]
```
The interval with highest average steps per day = `r maxStepsInterval`  


## Imputing missing values

The missing values will be resolved by imputing the previously computed mean  
steps per interval (not very fast code, sorry).
```{r}
print(nrow(activity) - sum(complete.cases(activity)))
newActivity <- activity
for (row in 1:nrow(newActivity)) {
    if (is.na(newActivity[row,]$steps)) {
        newActivity[row,]$steps <- stepsPerInterval[stepsPerInterval$interval == newActivity[row,]$interval,]$steps
    }
}

newStepsPerDay <- aggregate(steps~date, data = newActivity, sum)
hist(newStepsPerDay$steps)
newMeanSteps <- mean(newStepsPerDay$steps)
newMedianSteps <- median(newStepsPerDay$steps)

```
The average number of steps per day when NA values are resolved = `r meanSteps`  
The median number of steps per day when NA values are resolved = `r medianSteps`  


## Are there differences in activity patterns between weekdays and weekends?

```{r}
diffMean <- newMeanSteps - meanSteps
diffMedian <- newMedianSteps - medianSteps
```
The difference in average number of steps per day without and with NA values resolved = `r diffMean`  
The difference in median number of steps per day without and with NA values resolved = `r diffMedian`  

As can be seen in the following plots, the pattern for weekdays and weekends are quite different:
```{r, fig.height = 10}
Sys.setlocale("LC_TIME", "US")
newActivity$day <- weekdays(newActivity$date)
noWeekend <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
weekend <- c("Saturday", "Sunday")
newActivity$dayType <- factor(c("weekend", "weekday"))
newActivity[newActivity$day %in% noWeekend,]$dayType <- as.factor("weekday")
newActivity[newActivity$day %in% weekend,]$dayType <- as.factor("weekend")
par(mfrow = c(2, 1))
plot(aggregate(steps~interval, data = newActivity[newActivity$dayType == "weekday",], mean), type = "l")
plot(aggregate(steps~interval, data = newActivity[newActivity$dayType == "weekend",], mean), type = "l")
```



